### 2010-03-03 doc/msc-thesis/msc-nd.tex
initial msc thesis report

The Internet Draft \cite{draft-nd-07} specifies an optimized neighbor discovery mechanism sufficient for LoWPAN operation. The 6LoWPAN Neighbor Discovery (6LoWPAN-ND) introduces a node registration mechanism optimizing the node-router interface, which requires no flooding and reduces link-local multicast frequency.  The concept of a LoWPAN Whiteboard located at Edge Routers (routers that connect a LoWPAN to another IP network) is introduced, which allows for Duplicate Address Detection for the entire LoWPAN. The solution supports both Mesh Under and route over configurations for multihop forwarding. This chapter describes the specified 6LoWPAN-ND protocol.

@@

\section{Bootstrapping and Basic Operation}
Bootstrapping of a LoWPAN node consists of several steps.  At first, a node is required to autoconfigure at least one address, a link-local address, which is derived from the IEEE 64-bit extended MAC.  Therefore,  knowledge of the 64-bit address of a node is enough to derive its link-local address and reach it on the same link. In order to receive RAs from routers, a node joins the all-nodes multicast address and, if the node is a router, the all-routers multicast address. Once the interfaces have been initialized, a node listens for Router Advertisements (RA), or broadcasts a Router Solicitation (RS). Upon receipt of the RA, the node forms an optimistic global unique address with stateless address autoconfiguration and chooses one or more default routers (routers are present only in a route over configuration).

The constructed global address is tentative as long as the binding is not confirmed by the ER. To accomplish this, the node performs initial registration with a link-local ER  or LoWPAN Router by sending a unicast Node Registration (NR) message with a list of all addresses it wants to register. Registering directly with an ER is preferred, although all LoWPAN Routers have the ability to relay NR/NC messages on behalf of a node to the ER. The ER replies with a Node Confirmation(NC), which includes the set of addresses confirmed to be bound to the Whiteboard of the ER. Once the node has received the NC message, it is capable to send packets to any IPv6 address inside or outside the LoWPAN. Each router maintains a set of information about nodes that are currently registered through it, called the binding table. Nodes send periodic registration messages in order to maintain their bindings in the router binding tables and Edge Router Whiteboard. The detection of duplicate addresses (DAD) is also performed as part of the node registration process by the ER using a lookup on the Whiteboard.

@@

\subsection{Router Advertisement Message}
The format of RA message for 6LoWPAN is identical to the classic \cite{rfc4861} RA message. However, the use of a 2-bit  Default Router Preference (Prf) flag \cite{rfc4191} is defined in the context of 6LoWPAN. LoWPAN Edge Routers set Prf to $01$ indicating their high preference, whereas LoWPAN Routers with and without ER availability use the values $00$ and $01$ respectively for normal and low preference. RAs are sent either to all-nodes multicast or to a link-local unicast address as a response to an RS. RA messages use additional 6LoWPAN Information and 6LoWPAN Summary Options described in Sections \ref{nd.option.info} and \ref{nd.option.summary} respectively.

\subsection{Neighbor Solicitation/Advertisement Message}
Neighbor Solicitation(NS) and Neighbor Advertisement(NA) messages are employed only between ERs on the backbone link in Extended LoWPAN. Extended LoWPAN (shown in Figure \ref{fig:ext.lowpan})) allows to form a single subnet out of multiple LoWPANs interconnected by a backbone link via Edge Routers. The format of NS/NA messages is identical to the classic NS/NA messages \cite{rfc4861} with additional use of Owner Interface Identifier Option, described in Section \ref{sec:nd.oiio}.

@@

The message format for NR/NC messages is shown in Figure \ref{fig:nr.nc.format}. The values for the Type field will be defined in the future to distinguish Node Registration and Node Confirmation messages. The Status field is used only in NC messages and specifies the result of the registration procedure. A 4-bit unsigned Code provides additional details for the message, e.g. it may indicate whether the NR/NC is sent directly or relayed by a router. The Checksum field carries the ICMP checksum. A unique Transaction ID (TID) is used to match replies and a NC TID corresponds to the NR TID. Its value is incremented upon each new registration. A 1-bit Primary flag (P) indicates whether the ER is primary and may represent the node on the backbone in a Extended LoWPAN. A Router flag indicates the role of the node sending the NR message (0 is used by hosts and 1 by routers).  The Reserved field is unused and initialized to zero. The Binding Lifetime is the amount of time in minutes remaining before the binding of the owner interface identifier expires. The Advertising Lifetime field indicates the amount of time in units of 10 seconds the node will advertise itself to its local router using a NR local refresh. The Owner Nonce field is a 32-bit value generated randomly by the node upon booting.
And the Owner Interface Identifier (OII) is a globally unique identifier for the requesting host's interface.

@@

\subsubsection{6LoWPAN Address Option}\label{nd.option.address}
The 6LoWPAN Address Option (6AO) is used in NR and NC messages to indicate the address which a node wants to register with an ER and to get the result of registration back. A NR/NC message can include multiple Address Options. The fields of the option allow to indicate the success or failure of the binding in an NC, as well as to compress the carried IPv6 address to different extent.

@@

\section{Conceptual Data Structures }
\subsection{LoWPAN Node}
Every node in a LoWPAN keeps a list of prefixes advertised in Router Advertisements. The entries of the list are associated with the sequence numbers last advertised in the 6LoWPAN Summary Option, which allows to detect the changes in prefix information. In order to support stateful address compression, a node holds a list of context. As in the case of prefixes, each context entry is associated with the 6LoWPAN Summary Option sequence number. The list of Edge Routers the node is registered with an associated timeout and primary flag is stored within the Edge Router List. 
The Default Router List data structure contains the routers to which packets may be forwarded from the node.

In addition to the mentioned data structures, a node also keeps a set of conceptual variables for each interface, e.g. the default hop limit and the binding lifetime.

@@

\subsection{LoWPAN Edge Router}
Edge Routers implement LoWPAN Router features and extend them with Whiteboard registration for LoWPAN Nodes within their subnets. An ER maintains information about every registered node, e.g. the IPv6 address and Owner Interface Identifier of the node. There is no link-layer information stored in the Whiteboard. In an Extended LoWPAN the full registry of all the LoWPAN Nodes in a subnet is distributed between the Edge Routers.

>>>>

The internet draft \cite{draft-nd-07} specifies an optimized neighbor discovery mechanism sufficient for LoWPAN operation. The 6LoWPAN Neighbor Discovery (6LoWPAN-ND) introduces a node registration mechanism optimizing the node-router interface, which requires no flooding and reduces link-local multicast frequency.  The concept of a LoWPAN Whiteboard located at Edge Routers (ERs - routers that connect a LoWPAN to another IP network) is introduced, which allows for Duplicate Address Detection for the entire LoWPAN. The solution supports both Mesh Under and route over configurations for multihop forwarding. This chapter describes the specified 6LoWPAN-ND protocol.

@@

\section{Bootstrapping and Basic Operation}
Bootstrapping of a LoWPAN node consists of several steps.  At first, a node is required to autoconfigure at least one address, a link-local address, which is derived from the IEEE 64-bit extended MAC.  Therefore,  knowledge of the 64-bit address of a node is enough to derive its link-local address and reach it on the same link. In order to receive Router Advertisements (RAs) from routers, a node joins the all-nodes multicast address and, if the node is a router, the all-routers multicast address. Once the interfaces have been initialized, a node listens for RAs, or broadcasts a Router Solicitation (RS). Upon receipt of the RA, the node forms an optimistic global unique address with stateless address autoconfiguration and chooses one or more default routers (routers are present only in a route over configuration).

The constructed global address is tentative as long as the binding is not confirmed by the ER. To accomplish this, the node performs initial registration with a link-local ER  or LoWPAN Router by sending a unicast Node Registration (NR) message with a list of all addresses it wants to register. Registering directly with an ER is preferred, although all LoWPAN Routers have the ability to relay NR/NC messages on behalf of a node to the ER. The ER replies with a Node Confirmation(NC), which includes the set of addresses confirmed to be bound to the Whiteboard of the ER. Once the node has received the NC message, it is capable to send packets to any IPv6 address inside or outside the LoWPAN. Each router maintains information about nodes that are currently registered through it, called the binding table. Nodes send periodic registration messages in order to maintain their bindings in the router binding tables and Edge Router Whiteboard. The detection of duplicate addresses (DAD) is also performed as part of the node registration process by the ER using a lookup on the Whiteboard.

@@

\subsection{Router Advertisement Message}
The format of RA message for 6LoWPAN is identical to the classic \cite{rfc4861} RA message. However, the use of a 2-bit  Default Router Preference (Prf) flag \cite{rfc4191} is defined in the context of 6LoWPAN. LoWPAN Edge Routers set Prf to $01$ indicating their high preference, whereas LoWPAN Routers with and without ER availability use the values $00$ and $01$ respectively for normal and low preference. RAs are sent either to all-nodes multicast or to a link-local unicast address as a response to an RS. RA messages use additional 6LoWPAN Information and 6LoWPAN Summary Options described in Section \ref{nd.option.info}.

\subsection{Neighbor Solicitation/Advertisement Message}
Neighbor Solicitation(NS) and Neighbor Advertisement(NA) messages are employed only between ERs on the backbone link in an Extended LoWPAN. Extended LoWPAN (shown in Figure \ref{fig:ext.lowpan})) allows to form a single subnet out of multiple LoWPANs interconnected by a backbone link via Edge Routers. The format of NS/NA messages is identical to the classic NS/NA messages \cite{rfc4861} with additional use of the Owner Interface Identifier Option, described in Section \ref{sec:nd.oiio}.

@@

The message format for NR/NC messages is shown in Figure \ref{fig:nr.nc.format}. The values for the Type field will be defined in the future to distinguish Node Registration and Node Confirmation messages. The Status field is used only in NC messages and specifies the result of the registration procedure. A 4-bit unsigned Code provides additional details for the message, e.g., it may indicate whether the NR/NC is sent directly or relayed by a router. The Checksum field carries the ICMP checksum. A unique Transaction ID (TID) is used to match replies and a NC TID corresponds to the NR TID. Its value is incremented upon each new registration. A 1-bit Primary flag (P) indicates whether the ER is primary and may represent the node on the backbone in a Extended LoWPAN. A Router flag indicates the role of the node sending the NR message (0 is used by hosts and 1 by routers).  The Reserved field is unused and initialized to zero. The Binding Lifetime is the amount of time in minutes remaining before the binding of the owner interface identifier expires. The Advertising Lifetime field indicates the amount of time in units of 10 seconds the node will advertise itself to its local router using a NR local refresh. The Owner Nonce field is a 32-bit value generated randomly by the node upon booting.
And the Owner Interface Identifier (OII) is a globally unique identifier for the requesting host's interface.

@@

\subsubsection{6LoWPAN Address Option}\label{nd.option.address}
The 6LoWPAN Address Option (6AO) is used in NR and NC messages to indicate the address which a node wants to register with an ER and to get the result of the registration back. A NR/NC message can include multiple Address Options. The fields of the option allow to indicate the success or failure of the binding in an NC, as well as to compress the carried IPv6 address to different extent.

@@

\section{Conceptual Data Structures }
\subsection{LoWPAN Node}
Every node in a LoWPAN keeps a list of prefixes advertised in Router Advertisements. The entries of the list are associated with the sequence numbers last advertised in the 6LoWPAN Summary Option, which allows to detect the changes in prefix information. In order to support stateful address compression, a node holds a list of contexts. As in the case of prefixes, each context entry is associated with the 6LoWPAN Summary Option sequence number. The list of Edge Routers the node is registered with an associated timeout and primary flag is stored within the Edge Router List. 
The Default Router List data structure contains the routers to which packets may be forwarded from the node.

In addition to the mentioned data structures, a node also keeps a set of conceptual variables for each interface, e.g., the default hop limit and the binding lifetime.

@@

\subsection{LoWPAN Edge Router}
Edge Routers implement LoWPAN Router features and extend them with whiteboard registration for LoWPAN Nodes within their subnets. An ER maintains information about every registered node, e.g., the IPv6 address and Owner Interface Identifier of the node. There is no link-layer information stored in the Whiteboard. In an Extended LoWPAN the full registry of all the LoWPAN Nodes in a subnet is distributed between the Edge Routers.

### 2010-03-03 doc/msc-thesis/msc-routing.tex
initial msc thesis report

Table \ref{table:routing.prot.survey} summarizes the survey showing which of existing protocols meet the criteria described above. The detailed analysis is given in the document \cite{draft-protocols-07}. For each of these criteria, the value "pass" indicates that a protocol has satisfactory performance.  The value "fail" corresponds to not acceptable performance, which means that the protocol does not meet the criterion. Finally, a "?" means that a protocol would require a supplementary document specifying how a protocol should behave. As can be seen from the table, no existing IETF protocol meets all described criteria. Therefore, the survey concludes that new protocol specification documents have to be defined for a LLN routing protocol.

@@

\section{Overview of IPv6 Routing Protocol for LLNs}\label{sec:rout.rpl}
The IPv6 Routing Protocol for Low power and Lossy Networks (RPL) introduces a notion of a Directed Acyclic Graph (DAG), which is a directed graph with no cycles exist. All edges of a DAG are oriented toward and terminating at one or more root nodes uniquely identified by the DAGID. A Destination Oriented DAG (DODAG) is a DAG rooted at a single destination. While constructing a DODAG various routing metrics and optimization objectives can be used to compute the shortest paths. The set of supported link and node metrics used in the RPL is specified in the internet draft \cite{draft-routing-metrics-04}.

A DAG instance is a set of possibly multiple DODAGs. The routing metrics and objectives for a certain DAG instance are specified by an Objective Function (OF). Each DAG instance constructs a routing topology optimized for its OF. Traffic that belongs to a specific DODAG instance is marked in the flow label of the IPv6 header. A network may have more than one DAG instance which operate independently. This allows to applications to tag traffic to follow an appropriate DAG instance, i.e., optimized for low latency or low energy.

@@

\subsection{Basic Operation}
A node's neighbor set is the nodes that it can reach with a link-local multicast. A node that is not a DODAG root may maintain multiple DAG parents for a single DAG instance. DAG parents of a node is a subset of the set of       its neighbors. A node may implement different policies for selecting its DAG parents.

>>>>

Table \ref{table:routing.prot.survey} summarizes the survey showing which of existing protocols meet the criteria described above. The detailed analysis is given in the document \cite{draft-protocols-07}. For each of these criteria, the value "pass" indicates that a protocol has satisfactory performance.  The value "fail" corresponds to not acceptable performance, which means that the protocol does not meet the criterion. Finally, a question mark"?" means that a protocol would require a supplementary document specifying how a protocol should behave. As can be seen from the table, no existing IETF protocol meets all described criteria. Therefore, the survey concludes that new protocol specification documents have to be defined for a LLN routing protocol.

@@

\section{Overview of IPv6 Routing Protocol for LLNs}\label{sec:rout.rpl}
Typically, the traffic patterns for LLNs are not simply unicast, but in many cases point-to-multipoint or multipoint-to-point. The IPv6 Routing Protocol for Low power and Lossy Networks (RPL) \cite{draft-rpl-04} is designed to meet the routing requirements for LLNs and supports multipoint-to-point traffic from devices inside the LLN towards a central control point, as well as point-to-multipoint traffic from the central control point to the devices inside the LLN.

The RPL introduces a notion of a Directed Acyclic Graph (DAG), which is a directed graph with no cycles. All edges of a DAG are oriented toward and terminating at one or more root nodes uniquely identified by the DAGID. A Destination Oriented DAG (DODAG) is a DAG rooted at a single destination. A DODAG can be constructed with various routing metrics and optimization objectives in use to compute the shortest paths. The set of supported link and node metrics used in the RPL is specified in the internet draft \cite{draft-routing-metrics-04}.

A DAG instance may consist of multiple DODAGs. The routing metrics and objectives for a certain DAG instance are specified by an Objective Function (OF). Each DAG instance constructs a routing topology optimized for its OF. Traffic that belongs to a specific DODAG instance is marked in the flow label of the IPv6 header. A network may have more than one DAG instance which operate independently. This allows to applications to tag traffic to follow an appropriate DAG instance, i.e., optimized for low latency or low energy.

@@

\subsection{Basic Operation}
A set of nodes that can be reached with a link-local multicast forms the node's neighbor set. A node that is not a DODAG root may maintain multiple DAG parents for a single DAG instance. DAG parents of a node is a subset of the set of its neighbors and can be selected using different policies.
